### 2014-10-27 intro.tex
intro and motive

\section{Introduction}

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications.  While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web services, which aim
to provide an "always on" experience, overwhelmingly favour availability and
partition tolerance over strong consistency. To this end, several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have been proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an
item appear in a shopping cart after it has been
removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
application and is witnessed by the user.  Ultimately, the developer must
decide the consistency level appropriate for a particular operation; this is
understandably an error-prone process requiring intricate knowledge of both
the application as well as the semantics and implementation of the
underlying data store, which typically have only informal
descriptions~\cite{}.  Nonetheless, picking the correct consistency level is
critical not only for correctness but also for scalability of the
application. While choosing a weaker consistency level than required may
introduce program errors and anomalies, choosing a stronger one than
necessary can negatively impact program scalability.

%--KC--
% The choice of correct consistency level is critical to the correctness and
% scalability of the application. Choosing a weaker consistency level than
% required introduces anomalies, while choosing a stronger than necessary level
% affects program scalability.
%\KC{Do we include the LWW-Register number here to add emphasis to the fact that
%choosing the correct consistency level is essential for program scalability}.

Weak consistency also hinders compositional reasoning about programs.  While
an application might be naturally expressed in terms of well-understood and
expressive data types such as maps, trees, queues, or graphs,
geo-distributed stores typically only provide a minimal set of data types
with in-built conflict resolution strategies such as last-writer-wins (LWW)
registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while
traditional database systems enable composability through transactions,
geo-distributed stores typically lack unrestricted transactional access to
the data.  Working in this environment thus requires application state to be
suitably coerced to function using only the capabilities of the store.

To address these issues, we describe \name, a declarative programming model
and implementation for eventually consistent geo-distributed data
stores. The key novelty of \name is an expressive \emph{contract language}
to declare and verify fine-grained application-level consistency
properties. The programmer uses the contract language to axiomatically
specify the set of legal executions allowed over the replicated data
type. Contracts are constructed using primitive consistency relations such
as \emph{visibility} and \emph{session order} along with standard logical
and relational operators. A \emph{contract enforcement system} automatically
maps operations over the datatype to a particular consistency level
available on the store, and provably validates the correctness of the
mapping.  The paper makes the following contributions:

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a language for declaratively describing eventually
	consistent programs that manipulate replicated data types. Contracts are used
	to specify fine-grained application-level consistency properties, and are
	analyzed to assign the most efficient and sound store consistency level to
	the corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract enforcement
	system, and ensures that an operation is only executed if the required
	conditions on consistency are met.
\item \name is implemented as a transparent shim layer over
  Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
  extensive experimental evaluation over a collection of real-world
  applications, including a Twitter-like micro-blogging site and an
  eBay-like auction site illustrates the practicality of our approach.
\end{itemize}

The rest of the paper is organized as follows. Section~\ref{sec:motivation}
motivates \name and introduces the language through a series of examples.
Section~\ref{sec:lang} describes the contract language, and
Section~\ref{sec:opsem} presents an operational semantics.
Section~\ref{sec:txns} introduces transaction contracts and classification.
Section~\ref{sec:impl} describes the implementation and goes into detail about
the optimizations needed to make the system practical.
Section~\ref{sec:results} discusses experimental evaluation.
Section~\ref{sec:related} and~\ref{sec:concl} present related work and
conclusions.