### 2014-10-28 core-lang.tex

\section{Meta-theory}
\label{sec:core-opsem}

In this section, we present the meta-theoretic result that certifies
the soundness of our classification-based contract enforcement
strategy. 
% We show that any data store can correctly enforce $\scc$,
% $\ccc$, and $\ecc$ guarantees can satisfy application-level
% consistency constraints of all \name contracts.

Fig.~\ref{sem:oper} summarizes the artifacts needed to state our
meta-theoretic result. We formalize an abstract model of a replicated
data store as a tuple $\E;\Sigma$, where $\E$ denotes the execution
state, and $\Sigma$, the session soup, is the set of concurrent client
sessions interacting with the store. The execution state ($\E$) of a
data store is modeled as a tuple $\Exec$ where $\EffSoup$ is a set of
effects during the execution, and $\visZ,\soZ \subseteq \EffSoup
\times \EffSoup$ are visibility and session order relations,
respectively, witnessed over generated effects. A session is modeled
simply as a sequence of replicated data type operations ($op$), tagged
with the consistency class ($\tau$) of their contracts (as determined
by the contract classification scheme). We abbreviate contract classes
($\tau$) as uppercase {\sf SC}, {\sf CC}, and {\sf EC} for brevity.
Their representative contracts ($\cv(\tau)$) are, as usual, written
$\scc$, $\ccc$, and $\ecc$, respectively.  Our contract language and
classification scheme is independent of store semantics. Accordingly,
we abstain from making any specific assumptions about semantics of the
store. However, we do assume the existence of a reduction relation of
form:
\begin{smathpar} 
  \auxred{} {(\E,\Sigma)} {\eff} {(\E',\Sigma')}
\end{smathpar}
\noindent The relation captures store's progress in execution (from
$\E$ to $\E'$) due to the application of an operation from one of the
sessions in $\Sigma$ to a replica, generating a new effect ($\eff$).

Note that an execution state ($\E$) provides interpretations for
primitive relations (eg: $\visZ$) that occur free in contract
formulas, and also fixes the domain of quantification to set of all
effects ($\EffSoup$) observed during the execution. Therefore, $\E$ is
a potential model for any contract formula ($\cv$). If $\E$ is indeed
a model of $\cv$ (i.e., $\E \models \cv$), we say that the store
enforced the contract $\cv$ in execution $\E$.

% Core Operational Semantics
% --------------------------
\input{core-opsem.tex}

We now state the soundness of contract enforcement as following:

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be the contract of a replicated data type operation $op$,
and let $\tau$ denote the consistency class of $\cv$ 
as determined by the contract classification scheme. Forall execution
states $\E$, $\E'$ such that 
$\auxred{} {(\E,\langle (op,\tau);\,\sigma \rangle \pll \Sigma)} {\eff} 
 {(\E',\langle \sigma \rangle \pll \Sigma)}$
if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$ 
\end{theorem}

The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ guarantees in all executions, then contract
classification scheme can be used to satisfy any \name contract on
that data store. The proof of the theorem has been included in our
tech report~\cite{techrep}. Tech report also contains operational
semantics of our implementation of the data store, which concretizes
the reduction relation ($\xhookrightarrow{}$), along with the proof
that it correctly enforces all \name contracts.