### 2014-11-05 concl.tex
edits

\section{Conclusions}
\label{sec:concl}

In this paper, we have presented \name a shallow Haskell extension for
declarative programming over eventually consistent data stores. The key idea of
\name is the automatic classification of fine-grained consistency contracts on
operations and distributed transactions with respect to the consistency and
isolation levels offered by the store. Our contract language is carefully
crafted from a decidable subset of first-order logic, enabling the use of
automatic theorem prover to discharge the proof obligations associated with
contract classification. We provide meta-theory that certifies the correctness
of contract classification. We realize an instantiation of \name on top of
off-the-shelf distributed store, Cassandra, and illustrate the benefit of
fine-grained contract classification by implementing and evaluating several
scalable applications.