### 2014-11-05 motivation.tex

\section{Motivation}
\label{sec:motivation}

Consider how we might implement a highly available bank account on top of an
eventually consistent data store, with the \emph{integrity} constraint that the
balance must be non-negative. We begin by implementing a bank account
replicated data type (RDT) in \name, and then describe the mechanisms to obtain
the desired correctness guarantees.

\subsection{RDT specification}

A key novelty in \name is that it allows the addition of new RDTs to the store,
which obviates the need for coercing the application logic to utilize the store
provided data types. In addition, \name treats the convergence semantics of the
data type separately from its consistency properties. This separation of
concerns permits \emph{operational} reasoning for conflict resolution, and
\emph{declarative} reasoning for consistency. The combination of these
techniques enhances the programmability of the store.

Let us assume that the bank account object provides three operations:
\cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the assumption that the
withdraw fails if the account has insufficient balance. Every operation in
\name is of the following type, written in Haskell syntax:

\begin{codehaskell}
type Operation = [e] -> a -> (r, Maybe e)
\end{codehaskell}

\noindent It takes a list of effects (the \emph{context} for the operation),
and an input argument, and returns a result along with an optional effect
(read-only operations return \cf{Nothing}). The new effect (if emitted) is
added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}:

\begin{figure}
\begin{codehaskell}
data Acc = Deposit Int | Withdraw Int
				 | GetBalance

getBalance :: [Acc] -> () -> (Int, Maybe Acc)
getBalance ctxt _ =
  let res = sum [x | Deposit x <- ctxt]
						- sum [x | Withdraw x <- ctxt]
	in (res, Nothing)

deposit :: [Acc] -> Int -> ((), Maybe Acc)
deposit _ amt = (amt, Just dollar Deposit amt)

withdraw :: [Acc] -> Int -> (Bool, Maybe Acc)
withdraw ctxt v =
	if sel1 dollar getBalance ctxt () >= v
  then (True, Just dollar Withdraw v)
	else (False, Nothing)
\end{codehaskell}
\caption{Definition of a bank account expressed in Quelea.}
\label{fig:ex}
\end{figure}

The datatype \cf{Acc} represents the effect type for the bank account. The
context of the operations is a snapshot of the state of the object at some
replica. In this sense, every operation on the RDT is atomic, and thus
permitting sequential reasoning for implementing eventually consistent data
types. We have implemented a large corpus of RDTs for realistic benchmarks
including shopping carts, auction and micro-blogging sites in few tens of lines
of code.

\subsection{Anomalies under Eventual Consistency}

Our goal is to choose the correct consistency level for each of the bank
account operations such that (1) the balance remains non-negative and (2) the
\cf{getBalance} operation never incorrectly returns a negative balance. Let us
first consider the anomalies that could arise under eventual consistency.

\begin{figure}
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:cleanliness_examples}
\end{figure}

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with the
initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed
by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
witnesses the deposit and succeeds\footnote{Although visibility and session
order relations relate effects, we have abused the notation in these examples
to relate operations, with the idea that the relations relate the effect
emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \cf{getBalance} operation, that happens to witness all
the previous operations, would report a negative balance.

It is easy to see that preventing concurrent \cf{withdraw} operations
eliminates this anamoly. This can be done by insisting that \cf{withdraw} be
executed as a strongly consistent operation. Despite this strengthening,
\cf{getBalance} operation may incorrectly report a negative balance to the end
user. Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly},
which consists of three concurrent sessions performing a \cf{deposit}, a
\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
account object. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in
session 2, witnesses the effects of \cf{deposit(100)} from session 1, concludes
that there is sufficient balance, and completes successfully. However, the
\cf{getBalance} operation may only witness this successful withdraw, but not
the causally preceding \cf{dopsit}, and reports the balance of negative 50 to
the user.

Under eventual consistency, the users may also be exposed to other forms of
inconsistencies. Figure~\ref{fig:missingUpdateAnomaly} shows an execution where
the \cf{getBalance} operation in a session does not witness the effects of an
earlier \cf{withdraw} operation performed in the same session, possibly because
it was served by a replica that has not yet merged the \cf{withdraw} effect.
This anomaly leads the user to incorrectly conclude that the \cf{withdraw}
operation failed to go through.

Although it is easy to understand the reasons behind the occurrence of the
aforementioned anomalies, finding the appropriate fixes is not readily
apparent. Making \cf{getBalance} a strongly consistent operation is definitely
sufficient to avert anomalies, but is it really necessary? Given the cost of
enforcing strong consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid
it unless there are no viable alternatives. Exploring the space of these
alternatives requires understanding the subtle differences in semantics of
various kinds of weak consistency alternatives.

\subsection{Contracts}

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}\footnote{\name
exposes the contract construction language as a Haskell library}) that
axiomatically specify the set of allowed executions involving this operation.
In the case of the bank account, any execution that does not exhibit the
anomalies described in the previous section is a \emph{well-formed} execution
on the bank account object. By specifying the set of legal executions for each
data type in terms of a trace of operation invocations on that type, \name\
ensures that all executions over that type are well-formed.

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
vice versa. We express this application-level consistency requirement as a
contract ($\cv_w$) over \cf{withdraw}:

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
\end{array}
\end{smathpar}

\noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
The syntax $a:\rcf{withdraw}$ states that $a$ is an effect  emitted
by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
contract specifies that if the current operation emits an effect $\cureff$,
then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
Any execution on a bank account object that preserves the above contract for a
\cf{withdraw} operation is said to be derived from a correct implementation of
\cf{withdraw}.

\noindent For \cf{getBalance}, we construct the following contract ($\cv_{gb}$):

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
\end{array}
\end{smathpar}

\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
all \cf{deposit} operations $a$ visible to $b$ should also be visible to
$\cureff$. This prevents negative balance anomalies. Our contract language
provides operators to compose relations. The syntax $(R_1 \cap R_2)(a,b)$ is
equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the above contract
says that if a \cf{deposit} or a \cf{withdraw} operation precedes a
\cf{getBalance} operation in session order, and is applied on the same object
as the \cf{getBalance} operation, then it must be the case that the
\cf{getBalance} operation witnesses the effects of the preceding operations.

Finally, since there are no restrictions on when or how a \cf{deposit}
operation can execute, its contract is simply $\small \true$.

\subsection{From Contracts to Implementation}

Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
application-level consistency requirements, and make no reference to the
semantics of the underlying store. To write contracts, a programmer only needs
to reason about the semantics of the application under the \name system model.
The mapping of application-level consistency requirements to appropriate
store-level guarantees is done automatically behind-the-scene. How might one go
about ensuring that an execution adheres to a contract? The challenge is that a
contract provides a declarative (axiomatic) specification of an execution,
while what is required is an operational procedure for \emph{enforcing} its
implicit constraints.

One strategy would be to execute operations speculatively.  Here, operations
are tentatively applied as they are received from the client or other replicas.
We can maintain a runtime manifestation of executions, and check
well-formedness conditions at runtime, rolling back executions if they are
ill-formed. However, the overhead of state maintenance and the complexity of
user-defined contracts is likely to make this technique infeasible in practice.

We devise a static approach instead. Contracts are analyzed with the help of a
theorem prover, and statically mapped to a particular store-level consistency
property that the prover guarantees preserves contract semantics. We call this
procedure \emph{contract classification}. Given the variety and complexity of
store level consistency properties, the idea is that the system implementor
parameterizes the classification procedure by describing the store semantics in
the \emph{same} contract language as the one used to express the contract on
the operations. In the next section, we describe the contract language in
detail and describe the classification procedure for a particular store
semantics.