### 2014-11-06 evaluation.tex
edits

\section{Evaluation}
\label{sec:results}

In this section, we evaluate \name programs, report their contract profile and
illustrate the performance benefits of fine-grained consistency classification
on operations and transactions. We also evaluate on the impact of the
summarization. We implemented the following applications, which includes
individual RDTs as well as larger applications composed of several RDTs:

\begin{itemize}
\setlength{\itemsep}{2pt}
\item \textbf{LWW register}: A last-write-wins register that provides read and
write operations. Each write is associated with a timestamp, which is used to
resolve conflicting concurrent writes -- newer write wins.

\item \textbf{DynamoDB register}: A integer register that allows eventual and
strong puts and gets, conditional puts, increment and decrement operations.

\item \textbf{Bank account}: Our running example, with savings and current
accounts.

\item \textbf{Shopping list}: Collaborative shopping list which allows adding
and deleting items.

\item \textbf{Online store}: Models an online store with shopping cart and
dynamically changing item prices. Checkout process verifies that the customer
only pays the accepted price.

\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The application
allows users to browse items, bid for items on sale, and pay for items from a
wallet modelled after a bank account.

\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
Twissandra~\cite{Twissandra}. The application allows adding a new user, adding
and replying to tweets, following, unfollowing and blocking users, and fetching
a user's timeline, userline, followers and following.
\end{itemize}

\begin{table}
\setlength{\tabcolsep}{4pt}
{\sffamily \small
\begin{center}
\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
\hline
{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
\hline
{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
{RUBiS} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
\hline
\end{tabular}
\end{center} }
\caption{The distribution of classified contracts. \#T refers to the number of
tables in the application. The columns 4-6 (7-9) represent operations
(transactions) assigned to this consistency (isolation) level.}
\label{tab:ctrts}
\end{table}

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. In \name, the contract
classification process is completely performed at compile time and has no
overheads at runtime. The proof obligations associated with contract
classification is discharged through the Z3 SMT Solver. Across our benchmarks,
classifying a contract took 11.5 milliseconds on average.

For our performance evaluation, we deploy \name applications in \emph{clusters},
where each cluster is composed of 5 fully replicated Cassandra replicas within
the same datacenter. We instantiate one shim layer node for every Cassandra
replica, and place it on the same VM as the Cassandra replica. Clients are
instantiated on the same data center as the store, and run transactions. We
deploy the each node in the cluster on \cf{c3.4xlarge} Amazon EC2 instances.
Our shim layer nodes are multi-threaded, and we allocate 8 CPUs (out of 16
available) for each shim layer node. The clients also run on \cf{c2.4xlarge}
instances. We call this \cf{1DC} configuration. For our geo-distributed
experiments (\cf{2DC}), we instantiate 2 clusters, each with 5 nodes, and place
the clusters on US-east (Virginia) and US-west (Oregon). The average
inter-region latency was 85ms.

\begin{figure*}
  \centering
  \subfigure[Bank account]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-tp-vs-lat.pdf}}
  \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
  \subfigure[RUBiS bidding mix]{\label{grf:rubis}\includegraphics[width=0.24\textwidth]{graphs/Rubis.pdf}}
  \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
	\caption{Quelea Performance.}
  \label{grf:LWW_perf}
\end{figure*}

Figure~\ref{grf:BA-tp-vs-lat} shows the throughput vs. latency of operations
in bank account example as we increase the number of clients in \cf{1DC}
configuration. Our client workload was generated using YCSB
benchmark~\cite{YCSB}. The benchmark uniformly chose from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. We increased the
number of clients from 128 to 1024, and each experiment ran for 180 seconds.

The lines marked EC and CC correspond to all operations being assigned EC and
CC consistency levels. These levels compromise correctness as withdraw has to
be and SC operation. The line SC corresponds to a configuration where all
operations are strongly consistent; this ensures application correctness. \name
corresponds to our implementation, which classifies operations based on their
contracts. Both \name and SC ensure correctness. However, with 512 clients,
\name implementation was within 41\% of latency and 18\% of throughput of EC,
whereas SC operations had 162\% higher latency and 52\% lower throughput than
EC operations. Observe that there is a point in each case after which the
latency increases while the throughput decreases. This indicates a point after
which the store becomes saturated with client requests.

In \cf{2DC} configuration (not-shown), the average latency of SC operations
with 512 clients increased by 9.4$\times$ due to the cost of geo-distributed
coordination, whereas \name operations were only 2.2$\times$ slower, mainly due
to the increased cost of \cf{withdraw} operations. Importantly, the latency of
\cf{getBalance} and \cf{deposit} remained almost the same. This illustrates the
benefit of fine-grained contract classification in \name.

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
under 1DC configuration. The YCSB workload was modified to issue 10 operations
per transaction, with the default 50:50 read:write mix. Each operation is
assumed to have eventual consistency. \cf{NoTxn} corresponds to a configuration
that does not use transactions. Compared to this RC is only 12\% shower in
terms of latency with 512 clients, where as RR is 2.3X slower. The difference
between RC and \cf{NoTxn} is due to the meta-data overhead of recording
transaction information in the object state. For RR transaction, the cost of
capturing and maintaining the snapshot in an RR transaction is the biggest
source of overhead.

We also compared (not shown) the performance of EC LWW operations directly
against Cassandra (our backing store), which uses last-writer-wins as the only
convergence semantics. While Cassandra provides no stronger-than-eventual
consistency properties, \name was within 30\%(20\%) of latency(throughput) of
Cassandra with 512 clients, illustrating that the programmers only have to pay
a minimal overhead for the expressive and stronger \name programming model.

Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
configuration against a single replica (NoRep) and strongly replicated
(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
15\% read-write interactions, which is representative of the auction workload.
Without replication, NoRep trivially provides strong consistency. However, this
deployment does not scale beyond 1750 operations per second. Strong replication
offers better throughput at the cost of greater latency due to inter-replica
coordination. \name deployment offers the benefit of replication, while only
paying the cost of coordination when necessary.

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
replica, with all the clients operating on the \emph{same} LWW register to
stress test the summarization mechanism. The shim layer cache (mem) of
operations is summarized every 64 updates, while the updates in the backing
store (disk) are summarized every 4096 updates. Each point in the graph
represents the average latency of the previous 1000 operations. Each experiment
is run for 60s.

The results show that without summarization, the average latency of operations
increase exponentially to almost 1 second, and only 13K operations were
performed in a minute. Since every operation has to reduce over the set of all
previous operations, with a ever growing set, the operations take increasingly
more time to complete. With summarization only in memory, the performance still
degrades due to the cost of fetching all previous updates from the backing
store into the shim layer. Fetching the latest updates is essential for SC
operations. With both summarizations enabled, we see that the latency does not
increase over time, and we were able to perform 67K operations. This graph
illustrates the importance and effectiveness of summarization in \name.