### 2014-11-06 implementation.tex

\section{Implementation}
\label{sec:impl}

\name is implemented as a shallow extension of GHC Haskell and runs on top of
Cassandra, an off-the-shelf eventually consistent distributed data (or backing)
store responsible for all data management issues (i.e., replication, fault
tolerance, availability, and convergence).  Template Haskell is used implement
static contract classification, and proof obligations are discharged with the
help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod} illustrates the
overall system architecture.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\columnwidth]{Figures/ImplModel}
\end{center}
\caption{Implementation Model.}
\label{fig:impl_mod}
\end{figure}

Replicated data types and the stronger consistency semantics are implemented
and enforced in the \emph{shim layer}. Our implementation supports eventual,
causal, and strong consistency for data type operations, and RC, MAV, and RR
semantics for transactions.  This functionality is implemented entirely on top
of the standard interface exposed by Cassandra. From an engineering
perspective, leveraging an off-the-shelf data store enables an implementation
comprising roughly only 2500 lines of Haskell code, which is packaged as a
library.

\subsection{Shim Layer}

Each new effect in \name is realized as a row insertion in Cassandra, and the
state of an object is the set of all corresponding rows. The shim layer
maintains a causally consistent in-memory snapshot of a subset of objects in
the backing store, by explicitly tracking dependencies introduced between
effects due to visibility, session and same transaction relations. Dependence
tracking is similar to the techniques presented in~\cite{BoltOn}
and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
tolerance, each shim layer node simply acts as a soft-state cache, do not
communicate directly with each other, and can safely be terminated at any
point. Similarly, new shim layer nodes can be spawned on demand.

The shim layer nodes periodically fetch updates from the backing store for
eventually consistent operations, and on-demand for causally consistent and
strongly consistent operations. Strongly consistent operations are performed
after obtaining exclusive leases on objects. The lease mechanism is implemented
with the help of Cassandra's support for conditional updates and expiring
columns. Cassandra does not provide general-purpose transactions. Since the
transaction guarantees provided by \name are
coordination-free~\cite{BailisHAT}, we realize efficient implementations by
explicitly tracking the dependencies between operations and transactions.
Importantly, the weaker isolation semantics of transactions in \name permit
transactions to be discharged if at least one shim layer node is reachable.

\subsection{Summarization}

The main challenge in realizing an efficient implementation of operation-based
replicated data types is that the state of the object defined in terms of its
set of effects grows with every effectful operation. If left unchecked,
performance slow down over time, until the shim layer memory or backing store
disk runs out of memory. Luckily, the state of the operation-based replicated
data type can often be summarized to an \emph{observably equivalent} smaller
state. For example, a last-writer-wins register with multiple updates can be
summarized to an equivalent register with only the last write. Similarly, a
bank account with a series of deposits and updates with balance $b$ is
equivalent to a bank account with a single deposit of $b$. A set with a
collection of add and remove operations is equivalent to a set with a series of
addition of live elements from the original set.

Since the semantics of summarization is specific to each data type, \name
expects a summarization function (\cf{summarize}) of type \cf{[e]->[e]} for
each RDT, where the result is smaller than the argument but is observably
equivalent. We utilize the \cf{summarize} function to summarize the object
state both in the shim layer node and the backing store, typically when the
number of effects on an object crosses a tunable threshold. Shim layer
summarization is straight-forward; a summarization thread takes the local lock
on the object, and replaces its state with the summarized state. The shim layer
node remains unavailable for that particular object during summarization
(usually a few milliseconds).

Performing summarization in the backing store is more complicated; the whole
process needs to be atomic from a client's perspective, but Cassandra does not
provide multi-row transactions. We have engineered and implemented a scalable
summarization mechanism for the backing that permits concurrent client
operations, but nonetheless prohibits these operations from witnessing
intermediate states of the summarization process.