### 2014-11-07 concl.tex
edits

\section{Conclusions}
\label{sec:concl}

This paper presents \name, a shallow Haskell extension for declarative
programming over eventually consistent data stores. The key idea underlying
\name's design is the automatic classification of fine-grained consistency
contracts on operations and distributed transactions with respect to the
consistency and isolation levels offered by the store. Our contract language
is carefully crafted from a decidable subset of first-order logic, enabling
the use of automated verification tools to discharge the proof obligations
associated with contract classification. We realize an instantiation of
\name on top of off-the-shelf distributed store, Cassandra, and illustrate
the benefit of fine-grained contract classification by implementing and
evaluating several scalable applications.