### 2015-01-20 PLDI15_Response/response.txt
Edit rebuttal

We would like to thank the reviewers for their insightful comments. We first
address the concerns raised my multiple reviewers first followed by individual
concerns.

Complexity of contracts
-----------------------
We tend to disagree with the idea that the difficulty of expressing the
contracts is the same as choosing the correct store consistency level. While
the programmer does need to translate the integrity constraints in the
application to consistency constraints, the advantage is that the programmers
write contracts for the abstract system model. This results in generic RDTs
that can be realized on different data stores (as our implementation
illustrates) with varied consistency lattices.

The Quelea contract language itself is built upon three primitive,
well-understood relations -- vis, so and sameobj, along with the usual logical
and relational operators. Moreover, in the presence of coordination-free
transactions with weaker isolation levels, the choice of correct consistency
level is non-trivial. Quelea incorporates transactions with a single additional
primitive relation (sametxn). We believe the apparent complexity of contracts can
be further reduced using syntactic sugar, similar to our relational operators.

Comparison to previous work
---------------------------
Our work varies significantly from the previous works Bailis2013b and
Burckhardt2014. <!-- While Quelea provides a generic framework to describe new RDTs
and verifiably compile them to existing data stores, Bailis2013b is concerned
with implementing stronger consistency levels on top of existing stores.
Bailis2013b is not concerned with language abstractions, but rather an
implementation strategy. Our implementation, while sharing similarities with
Bailis2013b, illustrates that the Quelea programming model can be efficiently
realized on existing data stores. Moreover, Bailis2013b did not incorporate
transactions, which Quelea does. -->
<!-- GK: 1. I think the statement "a generic framework to describe new RDTs and
verifiably compile them to existing data stores" does not sufficiently
distinguish Quelea from Bailis2013b. -->

Bailis2013b describes implement strategies to realize stronger consistency
levels on top of eventually consistent stores. In contrast, Quelea provides a
language framework to describe application-level consistency requirements and
automatically map them to *some* implementation of appropriate store consistency
level. As such, Quelea's contract classification is independent of how various
consistency levels are actually implemented in the data store.  Furthermore,
Bailis2013b did not incorporate transactions, which Quelea does. Nevertheless,
our own implementation of strong consistency levels on top of Cassandra does
bear some similarities with Bailis2013b.

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
Quelea focuses on automatic verification of contracts using SMT solver, which
is essential for practicality. Quelea additionally incorporates support for
programming with coordination-free transactions.

Compile Time
------------
Our contract classification through the solver took on average took 11.5 ms
across the benchmarks (this statistic is also available on p.11 of the paper).
We expect this SMT solver overhead to remain low as the application scales. This
is because we do not analyze the RDT/transaction implementation to obtain the
verification conditions but rather simply analyze the user provided contracts.
Further, our contract language is defined carefully such that the validity check
is decidable for which efficient decision procedures exist. Moreover, our
contract language artifacts (e.g: primitive relations, effects), are encoded as
uninterpreted symbols in SMT without the need for expensive theories, such as
linear arithmetic.  Hence, the state space for the validity check remains low.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data introduced 4
incomparable session guarantees, and the consistency level for an operation is
any combination of the 4 session guarantees. The classification here, at worst,
requires 4! = 24 comparisons (SMT validity tests). Since each comparison is
very cheap, the solution remains scalable.

Review A
--------
The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to a
validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9 a,b,c
-- 5. Shim Layer nodes are soft-state entities and can be elastically scaled on
demand. All the benchmarks except Twissandra and RUBiS were written from
scratch.

Review B
--------
Bounded staleness could be accommodated into the contract language, but the
contract classification becomes undecidable since the theory of integer
arithmetic is undecidable.

RDT specification language does to some extent guards against
malicious/incorrect code. In particular, RDT operations are side-effect free.
While it is possible that the reduction can end up not terminating, it cannot
do any more harm.

A crashed operation does not affect the durable log. An operation is added to
the log only if it is successful. An operation is considered successful if it
produces a result.

R+ could be a total relation, but the R+ is still useful because it is not
required to be a total relation. For example, consider causal
consistency (CC) contract which is defined as:

 hbo \subseteq vis

where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
any superset of  transitive closure of (soo U vis). Nevertheless, all we can
assume about hbo is that it contains the  transitive closure of (soo U vis);
there is no proof of membership for other elements in hbo.  Consequently, by
asserting hbo \subseteq vis, the strongest guarantee that causal consistency 
contract is offering is that the transitive closure of (soo U vis) is contained
in vis. An application  contract (\psi) which requires more than the transitive
closure of (soo U vis) to be contained in vis  cannot be satisfied under CC
because the following implication is simply not valid even under our lenient
definition of “+“ :

 \Delta |- (hbo \subseteq vis) => \psi

\Delta, as usual, denotes the conjunction of basic axioms about primitive
relations.

Review E
--------
Our idea is to parameterize the classification procedure for each target data
store. Hence, if multiple partially ordered levels are appropriate, the
classification procedure can be encoded to choose the desired consistency
level based on external knowledge such as price.