### 2015-01-20 PLDI15_Response/response.txt
edits

We thank the reviewers for their detailed reviews and comments.

Complexity of contracts  (Reviewers ...)
-----------------------

We believe the difficulty of writing contracts is substantially simpler than
choosing the correct store consistency level since developers do not need to
understand the details of the underlying data store.

Moreover, while it is necessary to translate integrity constraints in the
application to consistency constraints as contracts, the tremendous
advantage that accrues is that these contracts are expressed over an
abstract system model, resulting in generic RDT specifications that can be
realized on different data stores (as our implementation illustrates) with
varied consistency lattices (as pointed out by ReviewerE).

The Quelea contract language itself is built upon *only* three primitive,
well-understood relations -- vis, so and sameobj, along with standard
propositional logical and relational operators.  In the presence of
coordination-free transactions with weaker isolation levels, the choice of
correct consistency level is non-trivial. Quelea incorporates transactions
with a single additional primitive relation (sametxn).

< provide an example using syntactic sugar >

Comparison to previous work (Reviewers ...)
---------------------------

Our work differs significantly from Bailis2013b and Burckhardt2014.

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike
Burckhardt2014, Quelea focuses on *automatic* verification of RDT
specifications, which is essential for practicality, and additionally
incorporates support for programming with coordination-free transactions.

Bailis2013b describes strategies to realize stronger consistency levels on
top of an eventually consistent store, but does not consider
parameterization of application consistency constraints over arbitrary data
store implementations.  In contrast, Quelea provides a language framework to
describe application-level consistency requirements and automatically map
them to *some* data store implementation. As such, Quelea's contract
classification is independent of how various consistency levels are actually
implemented in the data store.  Moreover, Bailis2013b does not incorporate
transactions, which Quelea does.

Compilation Time  (Reviewers ...)
----------------

Our contract classification through the solver took on average took 11.5 ms
across the benchmarks (this statistic is also available on p.11 of the
paper).  We expect this SMT solver overhead to remain low as the application
scales. This is because we do not analyze the RDT/transaction implementation
to obtain the verification conditions but only analyze user-provided
contracts.  Moreover, our contract language artifacts (e.g: primitive
relations, effects), are encoded as uninterpreted symbols in SMT without the
need for expensive theories, such as linear arithmetic.  Hence, the state
space for the validity check remains low.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data (cite ...)
introduced 4 incomparable session guarantees, and the consistency level for
an operation is any combination of the 4 session guarantees. The
classification here, at worst, requires 4! = 24 comparisons (SMT validity
tests). Since each comparison is very cheap, the solution remains scalable.

Review A
--------

The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to
a validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9
a,b,c -- 5. Shim Layer nodes are soft-state entities and can be elastically
scaled on demand. All the benchmarks except Twissandra and RUBiS were
written from scratch.

Review B
--------

Bounded staleness could be accommodated into the contract language, but
contract classification becomes undecidable since the theory of integer
arithmetic is undecidable.

The RDT specification language does to some extent guard against
malicious/incorrect code. In particular, RDT operations are side-effect
free.  While it is possible that the reduction can end up not terminating,
it cannot corrupt system state.

A crashed operation does not affect the durable log. An operation is added to
the log only if it is successful. An operation is considered successful if it
produces a result.

R+ could be a total relation, but R+ is still useful because it is not
required to be a total relation. For example, consider a causal
consistency (CC) contract defined as:

 hbo \subseteq vis

where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
any superset of the transitive closure of (soo U vis). Nevertheless, all we
can assume about hbo is that it contains the  transitive closure of (soo U
vis); there is no proof of membership for other elements in hbo. 
Consequently, by asserting hbo \subseteq vis, the strongest guarantee that
causal consistency  contract is offering is that the transitive closure of
(soo U vis) is contained in vis. An application  contract (\psi) which
requires more than the transitive closure of (soo U vis) to be contained in
vis  cannot be satisfied under CC because the following implication is
simply not valid even under our lenient definition of “+“ :

 \Delta |- (hbo \subseteq vis) => \psi

\Delta, as usual, denotes the conjunction of basic axioms about primitive
relations.

Review E
--------

Our idea is to parameterize the classification procedure for each target
data store. Hence, if multiple partially ordered levels are appropriate, the
classification procedure can be encoded to choose the desired consistency
level based on external knowledge such as price.