### 2015-01-21 PLDI15_Response/response.txt
edits

We thank the reviewers for their detailed reviews and comments.

Complexity of contracts  (Reviews B,C,D,E)
-----------------------

We believe the difficulty of writing contracts is substantially simpler than
choosing a correct store consistency level, since developers do not need to
understand the details of the underlying data store to define consistency
properties relevant for the application.

Moreover, while it is necessary to translate application-specific integrity
constraints to consistency constraints in the form of contracts, the tremendous
advantage that accrues is that these contracts are expressed over an abstract
system model, resulting in generic RDT specifications that can be realized on
different data stores (as our implementation illustrates) with varied
consistency lattices (as pointed out by ReviewerE).

The Quelea contract language itself is built upon *only* three primitive,
well-understood relations -- vis, so and sameobj, along with standard
propositional logical and relational operators.  In the presence of
coordination-free transactions with weaker isolation levels, the choice of
choosing the correct consistency level is non-trivial. Note also that Quelea
incorporates transactions with a single additional primitive relation
(sametxn).

Comparison to previous work (Reviews C,D)
---------------------------

Our work differs significantly from Bailis2013b and Burckhardt2014.

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
Quelea focuses on *automatic* verification of RDT specifications, which is
essential for practicality, and additionally incorporates support for
programming with coordination-free transactions.

Bailis2013b describes strategies to realize stronger consistency levels on top
of an eventually consistent store, but does not consider parameterization of
application consistency constraints over arbitrary data store implementations.
In contrast, Quelea provides a language framework to describe application-level
consistency requirements and automatically map them to *some* data store
implementation. As such, Quelea's contract classification is independent of how
various consistency levels are actually implemented in the data store.
Moreover, Bailis2013b does not incorporate transactions, which Quelea does.

Compilation Time  (Reviews A,D)
----------------

Our contract classification through the solver took on average took 11.5 ms
across the benchmarks (this statistic is also available on p.11 of the paper).
We expect this SMT solver overhead to remain low as the application scales.
This is because we do not analyze the RDT/transaction implementation to obtain
the verification conditions but only analyze user-provided contracts.
Moreover, our contract language artifacts (e.g: primitive relations, effects),
are encoded as uninterpreted symbols in SMT without the need for expensive
theories, such as linear arithmetic.  Hence, the state space for the validity
check remains low.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, [[Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data, PDIS '94]] introduced
4 incomparable session guarantees, and the consistency level for an operation
is any combination of the 4 session guarantees. The classification here, at
worst, requires 4! = 24 comparisons (SMT validity tests). Since each comparison
is very cheap, the solution remains scalable.

Review A
--------

The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to
a validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9
a,b,c -- 5. Shim Layer nodes are soft-state entities and can be elastically
scaled on demand. All the benchmarks except Twissandra and RUBiS were
written from scratch.

Review B
--------

Bounded staleness, which depends on wall clock time, is not incorporated in the
contract language; as integer arithmetic, in general, is undecidable, its
inclusion would make the classification procedure undecidable.

The standard benchmarks such as TPC-C are designed for traditional relational
database systems whose specifications dictate relational tables, joins and
transactions with strong isolation requirements, which are not readily
supported by NoSQL stores. As such these traditional benchmarks such as TPC are
not used to analyze and compare ECDS performance. On the other hand, since
Quelea was realized on top of Cassandra, we were able to obtain a point of
comparison with respect to vanilla Cassnadra (P.10, Column.2, para.3).

R+ could be a total relation, but R+ is still useful because it is not
*required* to be a total relation. For example, consider a causal
consistency (CC) contract defined as hbo \subseteq vis where hbo = (soo U
vis)+. Though our definition hbo can be a superset of transitive closure (TC)
of (soo U vis), we do not have a *proof* of inclusion of elements other than
the ones in the TC. As such an application contract (\psi) which requires more
than CC (say total order) would not be satisfied since the implication \Delta
|- (hbo \subseteq vis) => \psi is not valid.