### 2015-04-08 concl.tex

\section{Conclusions}
\label{sec:concl}

This paper presents \name, a shallow Haskell extension for declarative
programming over \ecds. The key idea underlying \name's design is the automatic
classification of fine-grained consistency contracts on operations and
distributed transactions with respect to the consistency and isolation levels
offered by the store. Our contract language is carefully crafted from a
decidable subset of first-order logic, enabling the use of automated
verification tools to discharge the proof obligations associated with contract
classification. We realize an instantiation of \name on top of off-the-shelf
distributed store, Cassandra, and illustrate the benefit of fine-grained
contract classification by implementing and evaluating several scalable
applications.