### 2015-04-10 related.tex

\section{Related Work}
\label{sec:related}

Operation-based RDTs have been widely studied in terms of their algorithmic
properties~\cite{SSS,Burckhardt2014}, and several systems utilize this model to
construct distributed data structures~\cite{Cassandra,Bayou,Tango}. These
systems typically propose to implement the datatypes directly over a cluster of
nodes, and only focus on basic eventual consistency. Hence, these systems
implement custom solutions for durability and fault-tolerance. \name realizes
stronger consistency models for RDTs on top of off-the-shelf eventually consistent
distributed stores. In this respect, \name is similar to~\cite{BoltOn} where
causal consistency is achieved through a shim layer layered on top of Cassandra, which
explicitly tracks and enforces dependencies between updates.
However,~\cite{BoltOn} does not support user-defined RDTs, automatic contract
classification and transactions.

Since eventual consistency alone is insufficient to build correct applications,
several systems~\cite{Bayou,Pileus,RedBlue} propose a lattice of stronger
consistency levels. Similarly, traditional database processing
systems~\cite{Berenson95} and their replicated variants~\cite{BailisHAT}
propose weaker isolation levels for performance. In these systems, the onus is
on the developer to choose the correct consistency(isolation) level for
operations(transactions). \name relieves the developer of this burden, and
instead only requires contracts to express declarative visibility requirements.

Our contract language and system model is inspired by the axiomatic
description of RDT semantics proposed in~\cite{Burckhardt2014}. While
~\cite{Burckhardt2014} uses axioms for formal verification of
an RDT implementation, we utilize them as a means for the user to express
desired consistency guarantees in the application. The operational semantics
of \name (described in the accompanying tech report~\cite{techrep}) describes
an operational manifestation of our system model. \cite{Burckhardt2015} also
presents an operational model of a replicated data store that is based on
the abstract system model presented in ~\cite{Burckhardt2014}. Their claims
about the expressivity and practicality of the system model vindicate
ours. However, both approaches differ in the way they aim to empower the application
programmer. \cite{Burckhardt2015} exposes the operational model to the
programmer, along with primitives, such as \cf{push}, \cf{pull}
and \cf{fetch}, that make it easy to \emph{implement} required
consistency and isolation guarantees. In contrast, \name completely
abstracts its operational model, and instead relies on contract
classification to automatically \emph{choose} appropriate consistency and isolation
levels. Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our
system model does not incorporate real (i.e., wall-clock) time. Hence, our
contract language cannot describe store semantics such as recency or
bounded-staleness guarantees offered by certain stores~\cite{Pileus}.

Several conditions have been proposed to judge whether an operation on a
replicated data object needs coordination or not. ~\cite{Calm} defines
\emph{logical monotonicity} as a sufficient condition for coordination
freedom, and proposes a consistency analysis that marks code regions
performing non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as
potential coordination points. ~\cite{IConfluence} and ~\cite{Sieve} define
\emph{invariant confluence} and \emph{invariant safety}, respectively, as
conditions for safely executing an operation without coordination.
~\cite{Sieve} also proposes a program analysis that conservatively marks
certain operations as \emph{blue} (coordination not required), while marking
the remaining as \emph{red} (coordination required). ~\cite{indigo} requires
programmers to declare application semantics, and the desired
application-specific invariants as formulas in first-order logic. It
performs static analysis on these formulas to determine $I$-offender sets -
sets of operations, which, when performed concurrently, result in violation
of one or more of the stated invariants. For each offending set of
operations, if the programmer chooses invariant-violation avoidance over
violation repair, the system employs various techniques, such as escrow
reservation, to ensure that the offending set is effectively
serialized. Unlike \name, these approaches focus on a coarse-grained
classification of consistency as eventual or strong, and do not consider
finer-grained transaction isolation levels. However, these analyses do
relieve programmers of the burden of tagging operations with consistency
levels. We consider automatic inference of consistency contracts from
application-specific integrity constraints as an obvious  next step for \name.