### 2015-04-15 evaluation.tex

\section{Evaluation}
\label{sec:results}

We present an evaluation study of our implementation, report contract
profiles of benchmark programs, and illustrate the performance benefits of
fine-grained consistency classification on operations and transactions. We
also evaluate the impact of the summarization. We have implemented the
following applications, which includes individual RDTs as well as larger
applications composed of several RDTs:

\begin{itemize}
\item \textbf{LWW register}: A last-write-wins register that provides read
	and write operations, where the read returns the value of the latest write.

\item \textbf{DynamoDB register}: An integer register that allows eventual
  and strong puts and gets, conditional puts, increment and decrement
  operations.

\item \textbf{Bank account}: Our running example.

\item \textbf{Shopping list}: A collaborative shopping list that allows
  concurrent addition and deletion of items.

\item \textbf{Online store}: An online store with shopping cart
  functionality and dynamically changing item prices.  The checkout process
  verifies that the customer only pays the accepted price.

\item \textbf{RUBiS}: An eBay-like auction site~\cite{RUBiS}. The
  application allows users to browse items, bid for items on sale, and pay
  for items from a wallet modeled after a bank account.

\item \textbf{Microblog}: A twitter-like microblogging site, modeled after
  Twissandra~\cite{Twissandra}. The application allows adding new users,
  adding and replying to tweets, following, unfollowing and blocking users,
  and fetching a user's timeline, userline, followers and following.
\end{itemize}

\begin{table}
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.2}
{\sffamily
\begin{center}
\caption{The distribution of classified contracts. \#T refers to the number of
tables in the application. The columns 4-6 (7-9) represent operations
(transactions) assigned to this consistency (isolation) level.}
\label{tab:ctrts}
\begin{tabular} {rrrrrrrrr}
\hline
{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
\hline
{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
{RUBiS} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
\hline
\end{tabular}
\end{center} }
\end{table}

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. In \name, the contract
classification process is completely performed at compile time and incurs no
runtime overhead. The proof obligations associated with contract
classification is discharged through the Z3 SMT Solver. Across our benchmarks,
classifying a contract took 11.5 milliseconds on average.

For our performance evaluation, we deploy \name applications in
\emph{clusters}, where each cluster is composed of five fully replicated
Cassandra replicas within the same datacenter. We instantiate one shim layer
node co-located on the same VM as a Cassandra replica. Clients are instantiated
within the same data center as the store, and run transactions. We deploy each
cluster and client node on a \cf{c3.4xlarge} Amazon EC2 instance. We call this
a \cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
instantiate 2 clusters, each with five nodes, and place the clusters on US-east
(Virginia) and US-west (Oregon) locations. The average inter-region latency was
85ms.

\begin{figure*}
  \centering
  \subfigure[Latency]{\label{grf:BA-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-lat.pdf}}
  \subfigure[Throughput]{\label{grf:BA-tp}\includegraphics[width=0.32\textwidth]{graphs/BA-tp.pdf}}
  \subfigure[Throughput vs. Latency]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-tp-vs-lat.pdf}}
	\caption{Bank account performance.}
  \label{grf:BA}
\end{figure*}

Figure~\ref{grf:BA} shows throughput vs. latency of operations in the bank
account example as we increase the number of clients in a \cf{1DC}
configuration. Our client workload was generated using the YCSB
benchmark~\cite{YCSB}. The benchmark uniformly chooses from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. We increased the
number of clients from 128 to 1024, and each experiment ran for 180 seconds.

The lines marked EC and CC correspond to all operations (including
\rcf{withdraw}) being assigned EC and CC consistency levels. These levels
compromise correctness as \rcf{withdraw} has to be an SC operation. The SC line
corresponds to a configuration where all operations are strongly consistent;
this ensures application correctness, at the cost of performance. \name
corresponds to our implementation, which classifies operations based on their
contract specifications. With 512 clients, the \name implementation was within
41\% of the latency and 18\% of the throughput of EC, whereas SC operations had
162\% higher latency and 52\% lower throughput than EC operations. Observe that
there is a point in each case after which the latency increases while the
throughput decreases; these correspond to points where the store becomes
saturated with client requests. In a \cf{2DC} configuration (not shown here),
the average latency of SC operations with 512 clients increased by 9.4$\times$
due to the cost of geo-distributed coordination, whereas \name operations were
only 2.2$\times$ slower, mainly due to the increased cost of \rcf{withdraw}
operations. Importantly, the latency of \rcf{getBalance} and \rcf{deposit}
remained almost the same, illustrating the benefit of fine-grained contract
classification.

\begin{figure*}
  \centering
  \subfigure[Latency]{\label{grf:LWW-txn-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-lat.pdf}}
  \subfigure[Throughput]{\label{grf:LWW-txn-tp}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp.pdf}}
  \subfigure[Throughput vs. Latency]{\label{grf:LWW-txn-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp-vs-lat.pdf}}
	\caption{LWW register transaction performance.}
  \label{grf:LWW-txn}
\end{figure*}

We compare the performance of different transaction isolation level choices in
Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
under a 1DC configuration. The YCSB workload was modified to issue 10
operations per transaction, with a default 50:50 read:write mix. Each operation
is assumed to be eventually consistent. NoTxn corresponds to a configuration
that does not use transactions. Compared to this, RC is only 12\% shower in
terms of latency with 512 clients, whereas RR is 2.3$\times$ slower. The
difference between RC and NoTxn is due to the meta-data overhead of recording
transaction information in the object state. For RR transactions, the cost of
capturing and maintaining a snapshot is the biggest source of overhead.

We also compared (not shown) the performance of EC LWW operations directly
against Cassandra (our backing store), which uses last-writer-wins as the only
convergence semantics. While Cassandra provides no stronger-than-eventual
consistency properties, \name was within 30\%(20\%) of latency(throughput) of
Cassandra with 512 clients, illustrating that the programmers only have to pay
a minimal overhead for the expressive and stronger \name programming model.

\begin{figure*}
  \centering
  \subfigure[Latency]{\label{grf:rubis-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-lat.pdf}}
  \subfigure[Throughput]{\label{grf:rubis-tp}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp.pdf}}
  \subfigure[Throughput vs. Latency]{\label{grf:Rubis-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp-vs-lat.pdf}}
	\caption{Rubis bidding mix performance.}
  \label{grf:rubis}
\end{figure*}

Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
configuration against a single replica (NoRep) and strongly replicated
(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
15\% read-write interactions, which is representative of the auction workload.
Without replication, NoRep trivially provides strong consistency. However, this
deployment does not scale beyond 1750 operations per second. Strong replication
offers better throughput at the cost of greater latency due to inter-replica
coordination. \name deployment offers the benefit of replication, while only
paying the cost of coordination when necessary.

\begin{figure}
\centering
\includegraphics[width=0.75\columnwidth]{graphs/summarization.pdf}
\caption{Impact of summarization.}
\label{grf:summarization}
\end{figure}

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We use 128 clients and a single \name replica,
with all clients operating on the \emph{same} LWW register to stress test the
summarization mechanism. The shim layer cache (memory) is summarized every 64
updates, while the updates in the backing store (disk) are summarized every
4096 updates. Each point in the graph represents the average latency of the
previous 1000 operations. Each experiment is run for one minute.  Without
summarization, the average latency of operations increases exponentially to
almost one second, and only 13K operations were performed in a minute. Since
every operation has to reduce over the set of all previous operations,
operations take increasingly more time to complete since they must contend with
an ever growing set.  With summarization only in memory, performance still
degrades due to the cost of fetching all previous updates from the backing
store into the shim layer. Fetching the latest updates from the backing store
is essential for SC operations. With summarization enabled on both disk and
memory, latency does not increase over time, and the implementation realizes
throughput of 67K operations/minute.